---
title: "Strawberry, Status, State, Toxicity"
author: "Group 1: Will, Zhihui, Shicong, Tao He"
date: "11/1/2021"
output: 
  revealjs::revealjs_presentation:
    theme: serif
    highlight: pygments
    self_contained: false
    reveal_options:
      slideNumber: true
      previewLinks: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA, echo = FALSE, message = FALSE, 
                      warning = FALSE, margin = FALSE)
pacman::p_load("ggplot2","revealjs","dplyr","tidyverse","magrittr",
               "hrbrthemes","viridis","corrplot","usmap","fmsb", "shiny")

library(shiny)
library(dplyr)
library(plotly)
library(shinydashboard)
library(RColorBrewer)
library(maps)
library(tigris)
library(leaflet)
library(stringr)
options(tigris_use_cache = TRUE)
```

# Introduction
<short>

- Organic/chemical data by state

- Which state used the most chemicals

- Toxicity to humans/bees
    + Carcinogen, hormone disruptor, neurotoxin, development
</short>


# Cleaning And Wrangle
<short>

- Cleaning the two data sets separately

- Normalization: use uppercase letter for value in chemical names

- Inner join the two data sets using chemical
</short>

## Details
<short>

<body style="color:DodgerBlue;">- Strawberry:</body> 
  - Remove meaningless columns
  - Separate Data.Item in order to get measurements
  - Remove , and (D), (Z), (NA) in value columns, convert it to numeric
  - Substitute space using underscore in chemical names

<body style="color:DodgerBlue;">- Pesticides:</body>
  - Fill the missing values with unknown
  - Add the chemicals that are in strawberries, not in pesticide
  - New column: toxicity level to bees/human, 0: no info, 1: harmful to bees/human
</short>

```{r}
straw <- read.csv("strawb.csv", header = TRUE)
```

## Date Frame


column names       | description
-------------------|------------------------------------------------------
chemical           | different chemical brands
chemical type      | different types of chemicals
toxicity-human/bee | whether the chemical has a toxicity to humans or bees
toxicity level     | "HIGH", "MODERATE", "SLIGHT", "UNKNOWN"...
state              | the states who plant strawberries
year               | from 2016 to 2019
measurements       | by using different units, like weight, sales,...



# Organic and Chemical
<short>
</short>

```{r}
# Chemical map
all_states <- map_data("state")  
 
p <- ggplot()
p <- p + geom_polygon(data=all_states, aes(x=long, y=lat, group = group),colour="gray", fill="#F4EDCA")

q <- filter(all_states, region == "california")
m<-  filter(all_states, region == "florida")
l<- filter(all_states, region == "oregon")
n<- filter(all_states, region == "washington")
s<- filter(all_states, region == "new york")
t<- filter(all_states, region == "north carolina")


par(mfrow=c(1,2))
p + geom_polygon(data = q, aes(x=long, y=lat, group = group),fill="#C3D7A4") +
    geom_polygon(data = m, aes(x=long, y=lat, group = group),fill="#56B4E9")+
    geom_polygon(data = l, aes(x=long, y=lat, group = group),fill="#CC99FF")+
    geom_polygon(data = n, aes(x=long, y=lat, group = group),fill="#FF9999")+
    labs(title="Map of strawberries with chemical")
```

## Organic Map

```{r}
p + geom_polygon(data = s, aes(x=long, y=lat, group = group),fill="#56B4E9")+
    geom_polygon(data = t, aes(x=long, y=lat, group = group),fill="#CC99FF")+
    labs(title="Map of organic strawberries")
```

# Chemical Usage 
<short>

- Select the chemical application measured in LB in the table.

</short>
```{r}
# density map 
measure_LB <- straw %>%
  filter(measurements == ' APPLICATIONS ,  MEASURED IN LB' & value>0)

measure_LB$value_log<- log(measure_LB$value)
plot_usmap(data = measure_LB, values = "value_log", color = "blue") + 
  scale_fill_continuous(low = "white", high = "blue", name = "Chemical Use by State", label = scales::comma) + 
  labs(title = "Chemical use ranking in four states", subtitle = "APPLICATIONS,MEASURED IN LB") +
  theme(legend.position = "right")
```

## Chemical Usage By Chemical Types
<short>

- Violin plots : witness the distribution of usage amount.

</short>

```{r}
# one for trend, one for distribution

# line chart, box plot
# violin chart
measure_LB <- straw %>%
  filter(measurements == ' APPLICATIONS ,  MEASURED IN LB' & value>0)

measure_LB %>%
  ggplot(aes(x=type, y=log(value), fill=type)) +
    geom_violin() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="B") +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("Violin chart") +
    xlab("")+
    ylab("Chemical Application,Measured in LB")
```

## Chemical Usage By Types In States

```{r}
z<- straw %>%
  count(Year, type, state)

ggplot(data = z,mapping = aes(x = Year, y = n, color=type)) +
  geom_point() +
  geom_smooth(method = 'loess',formula=y~x)+
  facet_grid(~state)
```


## Chemical Usage In Three Dimensions
<short>

- Let's combine three dimensions into one plot.

</short>


```{r}
measure_LB_year <- straw %>%
  filter(measurements == ' APPLICATIONS ,  MEASURED IN LB / ACRE / YEAR' & value>0)

ggplot(data = measure_LB_year,mapping = aes(x = as.factor(Year), y = log(value), color=type)) +
  geom_point() +
  geom_smooth()+
  xlab("Year")+
  ylab("Chemical Application,Measured in LB /ACRE/ YEAR") +
  facet_grid( type~state )

```

# Toxicity 

<short>- What are the factors that influence the choice of each state? 
  
  - Toxicity level
  
    - Humans(carcinogen, hormone disruption, Neurotoxins, 
           Developmental Or Reproductive Toxins)
  
    - Bees
    
  - Other factors
    
  - Question: Do people prefer to use the more toxic chemicals or the less toxic chemicals on different types of chemicals?</short>

## By Carcinogen

```{r}
par(mfrow = c(2,2))
straw_1 <- straw %>%
  filter(measurements %in% ' APPLICATIONS ,  MEASURED IN LB')

ggplot(straw_1) + 
  geom_bar(aes(x = type, fill = carcinogen), position = "fill")

```


## By Hormone Disruption

```{r}
ggplot(straw_1) + 
  geom_bar(aes(x = type, fill = hormone_disruptor), position = "fill")
```

## By Neurotoxins

```{r}
ggplot(straw_1) + 
  geom_bar(aes(x = type, fill = neurotoxins), position = "fill")
```


## By Developmental Or Reproductive Toxins

```{r}
ggplot(straw_1) + 
  geom_bar(aes(x = type, fill = developmental_or_reproductive_toxins), position = "fill")
```

# Toxicity To Bees 

```{r}
ggplot(straw_1) + 
  geom_bar(aes(x = type, fill = bee_toxins), position = "fill")
```

## The Whole Toxicity Level

<short>

- The area is the total toxicity level.

- In the insecticides people prefer to use more toxic chemicals.</short>
```{r}
# change the content of every toxicity level column
# build new variable
straw_1$Bee <- ifelse(straw_1$bee_toxins == "HIGH", 3, 
                    ifelse(straw_1$bee_toxins == "MODERATE", 2,
                           ifelse(straw_1$bee_toxins == "SLIGHT", 1, 0)))
straw_1$Carcinogen <- ifelse(straw_1$carcinogen== "KNOW", 3, 
                           ifelse(straw_1$carcinogen == "POSSIBLE", 2,
                                  ifelse(straw_1$carcinogen == "PROBABLE", 1, 0)))
straw_1$Hormone <- ifelse(straw_1$hormone_disruptor == "SUSPECTED", 1, 0)
straw_1$Developmental_or_reproductive_toxins <- ifelse(straw_1$developmental_or_reproductive_toxins == "PRESENT", 1, 0)
straw_1$Neurotoxins <- ifelse(straw_1$neurotoxins =="PRESENT", 1, 0)

# sum columns above
Type <- c("FUNGICIDE", "HERBICIDE", "INSECTICIDE", "OTHER" )
bee <- tapply(straw_1$Bee, straw_1$type, sum)
car <- tapply(straw_1$Carcinogen, straw_1$type, sum)
hor_disr <- tapply(straw_1$Hormone, straw_1$type, sum)
neu <- tapply(straw_1$Neurotoxins, straw_1$type, sum)
DRT <- tapply(straw_1$Developmental_or_reproductive_toxins, straw_1$type, sum)
# Custom the radar Chart
max_min <- data.frame(bee = c(180,0), car = c(180,0), hor_disr = c(180,0),
                      neu = c(180,0), DRT = c(180,0))
rownames(max_min) <- c("Max","Min")
data <- data.frame(bee, car, hor_disr, neu, DRT)
df <- rbind(max_min, data)
# plot a radar chart 
colors_border=c( rgb(0.1,0.5,0.5,0.9), rgb(0.8,0.3,0.5,0.9) , rgb(0.7,0.5,0.3,0.9) )
colors_in=c( rgb(0.1,0.3,0.5,0.7), rgb(0.6,0.1,0.2,0.4) , rgb(0.9,0.5,0.1,0.4) )
radarchart( df  , axistype=1 , 
    #custom polygon
    pcol=colors_border , pfcol=colors_in , plwd=4 , plty=1,
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(0,180,45), cglwd=0.8,
    #custom labels
    vlcex=0.8 
    )


# Add a legend
legend(x=1.2, y=1.2, legend = rownames(df[-c(1,4),]), bty = "n", pch=20 , col=colors_in , text.col = "grey", cex=1.2, pt.cex=3)
```

# Limitation And Improvement:
<short>
  
  - Missing data
    
  - No more additional data
  
  - Taking more data from website to explore this question later</short>

# Shiny Application

[Link for shiny](https://elisa99.shinyapps.io/visualization/)

# Conclusion
<short>
  
  - California > Florida > Oregon > Washington 
    
  - Further exploration of California’s fungicide and insecticide
  
  - Farms prefer more toxic</short>


